# Age Period Cohort analysis in R

Getting Data

```{r}
currentDataset <- read_csv("https://statsnotebook.io/blog/data_management/example_data/APC_cannabis.csv")
```

At level 1, we have

$logit P(cannabis_{ijk}) = b_0{jk}+b_1Age_{ijk}+b_2Age_{ijk}+b_3Female_{ijk}+b_4Inner regional_{ijk}+b_5Remote area_{ijk}+b_6High School_{ijk}$

where Female, Inner regional, Remote area and High School are indicator variables (i.e. High School = 0 for participants who haven’t finished high school and High School = 1 for participants who finished high school). 
R will automatically create these indicator variables when a categorical/factor variable is entered into a model. The quadratic term of age is included to capture the potential quadratic effect of age.

At level 2, we have

$b_0{jk} = γ_0+ u0_j + v0_k$

where $u0_j$ and $v0_k$ represents the effect of being in period j and birth cohort k, and $u0_j ~ N(0, τu)$ and $v0_j ~ N(0, τv)$. The $u0_j$ and $v0_k$ have also been referred to as the random effect of period and cohort. It should be noted that the index ijk is the index for individual i in period j and cohort k.


```{r}
#Converting Sex, HighSchool, Year, cohort_cat and remoteness into factor variable
currentDataset$Sex <- factor(currentDataset$Sex, exclude = c("", NA))
currentDataset$HighSchool <- factor(currentDataset$HighSchool, exclude = c("", NA))
currentDataset$Year <- factor(currentDataset$Year, exclude = c("", NA))
currentDataset$cohort_cat <- factor(currentDataset$cohort_cat, exclude = c("", NA))
currentDataset$remoteness <- factor(currentDataset$remoteness, exclude = c("", NA))

#centre age at 20 years old and create the quadratic term
currentDataset$AgeR = currentDataset$AgeR - 2
currentDataset$AgeR2 = currentDataset$AgeR^2

```


```{r}
install.packages('lme4')
```


```{r}
library(lme4)

res <- glmer(weekly_cannabis ~ AgeR + AgeR2 + Sex + remoteness + HighSchool + (1 | cohort_cat) + (1 | Year),
  family = binomial,
  data = currentDataset)
summary(res)
confint(res, level = 0.95, method = "Wald")

```
```{r}
se <- sqrt(diag(vcov(res)))
z <- -qnorm((1-0.95)/2)
exp(cbind(Est=fixef(res),
  "2.5%"  = fixef(res) - z*se,
  "97.5%" = fixef(res) + z*se))

```

As expected, age is strongly associated with reduced odds of using cannabis weekly. The quadratic term of age is not statistically significant and is very close to zero.

In the presence of the quadratic term of age and other variables in the model, to better understand the effect of age on cannabis use, we can use the **emmeans** package to calculate the probability of weekly cannabis use with the following codes.

```{r}
library(emmeans)
```


```{r}

Age_eff = data.frame()

# Age is centred at 20 years old, and is divided by 10.
# (i.e. 0 means the participant is 20 years old, 1 means 30 years old, etc)
# We use a for loop to calculate the probability of weekly cannabis use 
# when the age "score" is between 0 and 5, with 0.5 increment.
# The calculated probability is stored in a new data frame Age_eff

for (i in 0:10) {
  emm <- emmeans(res,  ~ AgeR + AgeR2,
  at = list(
  AgeR = i/2, AgeR2 = (i/2)^2), type = "response", level = 0.95)
  Age_eff = rbind(data.frame(summary(emm)), Age_eff)
}

```

The above codes prodce the following output. This can be considered as the “average” probability averaging across other variables. 

Note that a score of “0” for age represents 20 years old and a score of “5” represents 70 years old. The column prob is the estimated probability of weekly cannabis use averaged across levels of sex, remoteness, and high school completion

```{r}
head(Age_eff)
```

```{r}
library(ggplot2)
```

```{r}
#We transformed back the x axis from "age score" into actual age by adding 2 and then multiplying by 10.
plot <- ggplot(Age_eff, aes(x = ((AgeR+2)*10), y = prob)) +
    geom_point() + 
    geom_line() +
    geom_errorbar(aes(ymin = asymp.LCL, ymax = asymp.UCL), width = 1) +
    xlab("Age") +
    ylab("Probability of weeky cannabis use")
    
plot
```
```{r}
res_cohort_period <- res
```

```{r}
res <- glm(weekly_cannabis ~ AgeR + AgeR2 + Sex + remoteness + HighSchool,
  family = binomial,
  data = currentDataset)
summary(res)
confint(res, level = 0.95, method = "Wald")

```

# Testing Cohort effect

```{r}
anova(res_cohort_period, res)
```

# Testing period effect

```{r}
res <- glmer(weekly_cannabis ~ AgeR + AgeR2 + Sex + remoteness + HighSchool + (1 | cohort_cat),
  family = binomial,
  data = currentDataset)
summary(res)
confint(res, level = 0.95, method = "Wald")

res_cohort <- res

anova(res_cohort_period, res_cohort)

```


# Visualising Cohort effect

```{r}
#Extract the random effect
u0 <- ranef(res_cohort_period, condVar = TRUE)
names(u0$Year) <- "est"
names(u0$cohort_cat) <- "est"

#Extract the standard error
period_eff <- data.frame(est = u0$Year, se = sqrt(attr(u0[[1]], "postVar")[1, ,]),
  period = c(2001, 2004, 2007, 2010, 2013, 2016, 2019))
cohort_eff <- data.frame(est = u0$cohort_cat, se = sqrt(attr(u0[[2]], "postVar")[1, ,]),
  cohort = c("1941/50","1951/60","1961/70","1971/80","1981/90","1991/2000"))

period_eff$upper <- period_eff$est + 1.96*period_eff$se
period_eff$lower <- period_eff$est - 1.96*period_eff$se
cohort_eff$upper <- cohort_eff$est + 1.96*cohort_eff$se
cohort_eff$lower <- cohort_eff$est - 1.96*cohort_eff$se

#Visualise the period and cohort effect using ggplot2
plot <- ggplot(period_eff, aes(x = period, y = est)) +
  geom_point() + 
  geom_errorbar(aes(ymin = lower, ymax = upper), width = .2) + 
  xlab("Year") + 
  ylab("Conditional log odds of the period effect")

plot

plot <- ggplot(cohort_eff, aes(x = cohort, y = est)) +
  geom_point() + 
  geom_errorbar(aes(ymin = lower, ymax = upper), width = .2) + 
  xlab("Year") + 
  ylab("Conditional log odds of the cohort effect")

plot

```





















