---
title: "Lab 16"
authot: Ipshita Ghosh 20122006
output: pdf_document
---

# Text Analysis - Sentiment Analysis

```{r}
# load in the libraries we'll need
library(tidyverse)
library(tidytext)
library(glue)
library(stringr)
library(SnowballC)
library(tm)
library(twitteR)
library(syuzhet)
```

Read the file
```{r}
df <- read.csv('C:\\Users\\Dell\\OneDrive\\College_2nd\\R Lab\\text_data.csv')
```

Looking at the file
```{r}
glimpse(df)
```

Looking at the dimension of the data
```{r}
length(df)
```
```{r}
dim(df)
```

Head of the file
```{r}
head(df)
```

Renaming the file
```{r}
df <- rename(df, target = X0)
df <- rename(df, id = X1467810369)
df <- rename(df, date = Mon.Apr.06.22.19.45.PDT.2009)
df <- rename(df, query = NO_QUERY)
df <- rename(df, user = X_TheSpecialOne_)
df <- rename(df, text = X.switchfoot.http...twitpic.com.2y1zl...Awww..that.s.a.bummer...You.shoulda.got.David.Carr.of.Third.Day.to.do.it...D)
```


```{r}
head(df)
```

Cleaning the file
```{r}
#Removing hashtag , urls and other special characters

df$text <- gsub("http.*","",df$text)

df$text <- gsub("https.*","",df$text)

df$text <- gsub("#.*","",df$text)

df$text <- gsub("@.*","",df$text)

```


More cleaning
```{r}
clean_tweets <- function(x) 
  {
  x %>%
  str_remove_all(" ?(f|ht)(tp)(s?)(://)(.*)[.|/](.*)") %>%
  str_replace_all("&amp;", "and") %>%
  str_remove_all("[[:punct:]]") %>%
  str_remove_all("ˆRT:? ") %>%
  str_remove_all("@[[:alnum:]]+") %>%
  str_remove_all("#[[:alnum:]]+") %>%
  str_replace_all("\\\n", " ") %>%
  str_to_lower() %>%
  str_trim("both")
  return (x)
  }
# applying the cleaning function to tweet column
df$text = sapply(df$text, clean_tweets)

```

Getting sentiments
```{r}
get_sents = function(tweet)
  {
  sent.value = get_sentiment(tweet)
  return (sent.value)
  }
# applying the get_sents function to tweet column
df$sentiment_score = sapply(df$text, get_sents)
```


Looking at the dataframe
```{r}
head(df)
```


Assigning the tweets as positive or negative
```{r}
positive.tweets = df[df$sentiment_score > 0,]
negative.tweets = df[df$sentiment_score < 0,]
```


Assigning sentiment labels to tweets
```{r}
to_sent_labels = function(x)
  {
  if (x >0)
    {
  'positive'
  }
  else if (x <0)
    {
    'negative'
    }
  else
    {
'neutral'
    }
}

df$sent_labels = sapply(df$sentiment_score , to_sent_labels)
head(df)

```

```{r}
print(table(df$sent_labels))

```

Printing the sentiments

```{r}
library(ggplot2)
ggplot(data = df, aes(sent_labels)) + geom_bar(fill = 'lightgreen') + theme_light()
```

# Inference

We can see that most of the tweets are neutral, making up about 65% of the data. Positive and negative tweets are fairly equally divided.



















