# Decision tree in R

---
title: "Lab 13"
author: Ipshita Ghosh 20122006
---

Library for Decision tree
```{r}
library(rpart)
library(rpart.plot)
```

Library for plotting
```{r}
library(ggplot2)
```

Library to plot missing data
```{r}
library(naniar)
```

Library to tidy our dataset
```{r}
library(tidyr)
```

Getting the data
```{r}
data <- read.csv('C:\\Users\\Dell\\OneDrive\\College_2nd\\R Lab\\accident.csv')
```


# EDA and Preprocessing

Getting a look at the data
```{r}
head(data)
```

Looking at the structure of data
```{r}
str(data)
```

Statistical summary of data
```{r}
summary(data)
```

Visualising Missing data

```{r}
vis_miss(data)
```

Since the error percentage is very low, dropping the NA values
```{r}
df <- na.omit(data)
```

Looking at the graph again
```{r}
vis_miss(df)
```

Deleting unecessary columns

```{r}
df = subset(df,select= c(-id,-caseid))
```

<\hr> 

Making few columns of the dataframe as factor

```{r}
df$dead <- as.factor(df$dead)
df$airbag <- as.factor(df$airbag)
df$seatbelt <- as.factor(df$seatbelt)
df$frontal <- as.factor(df$frontal)
df$sex <- as.factor(df$sex)
df$abcat <- as.factor(df$abcat)
df$occRole <- as.factor(df$occRole)
df$deploy <- as.factor(df$deploy)
df$injSeverity <- as.factor(df$injSeverity)
```


Looking at the structure of data frame

```{r}
str(df)
```
<\hr>

Plotting a few Graphs

```{r}
ggplot(data,aes(dead))+
        geom_bar(aes(fill=factor(dead)),alpha=0.5)
```
AIRBAG

```{r}
ggplot(df,aes(airbag))+
        geom_bar(aes(fill=factor(airbag)),alpha=0.5)
```

SEATBELT

```{r}
ggplot(df,aes(seatbelt))+
        geom_bar(aes(fill=factor(seatbelt)),alpha=0.5)
```

FRONTAL

```{r}
ggplot(df,aes(frontal))+
        geom_bar(aes(fill=factor(frontal)),alpha=0.5)
```

SEX

```{r}
ggplot(df,aes(sex))+
        geom_bar(aes(fill=factor(sex)),alpha=0.5)
```

INJURE SEVERITY

```{r}
ggplot(df,aes(injSeverity))+
        geom_bar(aes(fill=factor(injSeverity)),alpha=0.5)
```

YEAR OF ACCIDENT

```{r}
plot(df$yearacc)
```

<\HR>

# Modelling

encoding our target value as 0 and 1

```{r}
to_encod = function (x)
  {
    if (x=='alive')
      {1}
  else{0}
}
```

Making a new column y and assigning dead to it

```{r}
df$y = sapply(df$dead,to_encod)
```

Dropping the dead column

```{r}
df = subset(df,select= c(-dead))
```

<\HR>

SPLITTING THE DATA

```{r}
create_train_test <- function(data, size = 0.8, train = TRUE)
{
  #' create_train_test(df, size = 0.8, train = TRUE)
  #' arguments:
  #' @param df: Dataset used to train the model.
  #' @param size: Size of the split. By default, 0.8. Numerical value
  #' @param train: If set to `TRUE`, the function creates the train set, otherwise the test set. Default value sets to `TRUE`. Boolean value.You      need to add a Boolean parameter because R does not allow to return two data frames simultaneously.
  
  #' @return test/train data
  
  
    n_row = nrow(data)
    total_row = size * n_row
    train_sample <- 1: total_row
    if (train == TRUE) {
        return (data[train_sample, ])
    } else {
        return (data[-train_sample, ])
    }
}
```

```{r}
data_train <- create_train_test(df, 0.8, train = TRUE)
data_test <- create_train_test(df, 0.8, train = FALSE)
```
Dimension of the train and test

```{r}
dim(data_test)
```
```{r}
dim(data_train)
```

Looking at the proportion of y in the train and test

```{r}
prop.table(table(data_test$y))
```
```{r}
prop.table(table(data_train$y))
```

<\hr>

Making the decision tree

```{r}
fit <- rpart(y~., data = data_train, method = 'class')
rpart.plot(fit)
```
Predict on unseen dataset

```{r}
predict_unseen <-predict(fit, data_test, type = 'class')
```

Looking at true positive and negatives

```{r}
table_mat <- table(data_test$y, predict_unseen)
table_mat
```

CHECKING ACCURACY OF THE MODEL

```{r}
accuracy_Test <- sum(diag(table_mat)) / sum(table_mat)
print(paste('Accuracy for test', accuracy_Test))
```