---
output:
  html_document: default
  pdf_document: default
---
# Lab 10
# 06th April 2021

Loading the Dataset

```{r}
setwd("C:\\Users\\Dell\\OneDrive\\College_2nd\\R Lab") 
df <- read.csv('Admission_Predict.csv')
```


Looking at the dataset
```{r}
head(df)
```

Descriptive statistical summary
```{r}
summary(df)
```

Structure of the dataframe
```{r}
#structure
str(df)
```
We can see that some of the variables are factors and some are numericals


Looking for missing values
```{r}
#null
any(is.na(df))
```
We dont have any missing values


## Plotting

For Research
```{r}
library(ggplot2)
ggplot(df,aes(Research))+
        geom_bar(aes(fill=factor(Research)),alpha=0.5)
```
We can see that the more people have researches but the difference is not wide.


```{r}
ggplot(df,aes(University.Rating))+
        geom_bar(aes(fill=factor(University.Rating)),alpha=0.5)
```
Most of the University are medium ranked, the distribution is slightly skewed



```{r}
ggplot(df,aes(GRE.Score))+geom_histogram(fill='blue', bins = 20,alpha=0.5)
```
The GRE scores dont seem to have a pattern


```{r}
ggplot(df,aes(TOEFL.Score))+geom_histogram(fill='blue', bins = 20,alpha=0.5)
```
The TOEFL score is somewhat similar to the GRE Score, with not visible pattern as such.


```{r}
ggplot(df,aes(CGPA))+geom_histogram(fill='blue', bins = 20,alpha=0.5)

```
The CGPA has no visible patterns

```{r}
ggplot(df,aes(SOP))+geom_histogram( bins = 20,alpha=0.5)
```
SOP follows a normal distribution with slight skew


```{r}
ggplot(df,aes(LOR))+geom_histogram( bins = 20,alpha=0.5)
```



```{r}
plot(df$Chance.of.Admit,xlab = 'Index', ylab = 'Chance of admission')
```
We can see that the Chance of Admit follows no such patters, hence we will create another column named Status. Status will have have the the value 0 if the probability of admission is less than 70% and 1 otherwise.

```{r}
library(dplyr)
```
```{r}
df <- df %>%
  mutate(Status = case_when(
    Chance.of.Admit>0.7 ~ 1,
    Chance.of.Admit<=0.7 ~ 0
    ))
```

```{r}
plot(df$Status, xlab = 'Index',ylab = 'Status 1/0')
```
This data can be fit into logistic model

```{r}
cor(df)
```
From the correlation table, choosing the important and highest correlated values. Chance of Admit is highly collinear with other values and status in itself is a representation of Chance of Admit, hence it will not be added


```{r}
glm.fit=glm(Status∼GRE.Score +  TOEFL.Score + University.Rating + SOP + LOR + CGPA +Research,
        data=df ,
        family=binomial )

summary (glm.fit)
```


```{r}
#predictions
glm.probs=predict (glm.fit ,type="response")
glm.probs [1:10]
```
```{r}
glm.pred=rep(0 ,400)

glm.pred[glm.probs >.7]= 1
```
In order to understand the result in the basis of 0 and 1, we need to convert the probabilities accordingly. Hence the values with more than 0.7 are marked 1 and those with less that 0.7 are marked 0



CONFUSION MATRIX
```{r}
table(glm.pred ,df$Status)
```
Hence our model correctly predicted that the admission will be done 180 times and that it would not be on 154 times. 
For a total of 154+180 = 334 correct predictions.

The mean() function can be used to compute the fraction of times for which the prediction was correct. In this case, logistic regression
correctly predicted the movement of the market 83.5 % of the time.

```{r}
mean(glm.pred==df$Status)
```
Sensitivity = TP / (TP + FN)
1 - Specificy = TN / (TN + FP)

```{r}
Sensitivity = 180/(180 + 55)
Specificity = 1 - (154/(154+11))

cat("Sensitivity is", Sensitivity, "and Specificity is", Specificity,"\n")
```


































