---
output:
  html_notebook: default
  html_document: default
---
# Decision Tree in R

Library for data wrangling
```{r}
library(dplyr)
```

Library for decision tree
```{r}
library(rpart)
library(rpart.plot)
```

Library for plotting
```{r}
library(ggplot2)
```


Importing the data
```{r}
data = read.csv('C:\\Users\\Dell\\OneDrive\\College_2nd\\R Lab\\adult.csv')
```

Looking at the data
```{r}
head(data)
```

Statistical summary

```{r}
summary(data)
```

Structure of the data
```{r}
str(data)
```

Checking for Nulls
```{r}
any(is.na(data))
```

Converting to factors:
```{r}
data$workclass <- factor(data$workclass, exclude = c("", NA))

data$education <- factor(data$education, exclude = c("", NA))

data$education.num <- factor(data$education.num, exclude = c("", NA))

data$marital.status <- factor(data$marital.status, exclude = c("", NA))

data$occupation <- factor(data$occupation, exclude = c("", NA))

data$relationship <- factor(data$relationship, exclude = c("", NA))

data$race <- factor(data$race, exclude = c("", NA))

data$sex <- factor(data$sex, exclude = c("", NA))

data$income <- factor(data$income, exclude = c("", NA))

data$hours.per.week <- factor(data$hours.per.week, exclude = c("", NA))

data$native.country <- factor(data$native.country, exclude = c("", NA))
```



Looking at structure again
```{r}
str(data)
```

Dropping unwanted columns
```{r}
# Drop variables
df <- data %>% select(-c(fnlwgt, capital.gain))
```


```{r}
glimpse(df)
```


```{r}
head(df)
```

Plotting Education

```{r}
ggplot(df,aes(education))+
        geom_bar(aes(fill=factor(education)),alpha=0.5)
```
We can see that most of them are high school graduates


Plotting Marital status
```{r}
ggplot(df,aes(marital.status))+
        geom_bar(aes(fill=factor(marital.status)),alpha=0.5)
```

Plotting Occupation
```{r}
ggplot(df,aes(occupation))+
        geom_bar(aes(fill=factor(occupation)),alpha=0.5)
```

Plotting Race
```{r}
ggplot(df,aes(race))+
        geom_bar(aes(fill=factor(race)),alpha=0.5)
```
The data is clearly unbalanced and biased


Plotting sex

```{r}
ggplot(df,aes(sex))+
        geom_bar(aes(fill=factor(sex)),alpha=0.5)
```
Again we can see that the classes are imbalanced


Plotting income
```{r}
ggplot(df,aes(income))+
        geom_bar(aes(fill=factor(income)),alpha=0.5)
```
The target class is highly imbalanced


## Modelling

Create test and train

```{r}
create_train_test <- function(data, size = 0.8, train = TRUE)
{
  #' create_train_test(df, size = 0.8, train = TRUE)
  #' arguments:
  #' @param df: Dataset used to train the model.
  #' @param size: Size of the split. By default, 0.8. Numerical value
  #' @param train: If set to `TRUE`, the function creates the train set, otherwise the test set. Default value sets to `TRUE`. Boolean value.You      need to add a Boolean parameter because R does not allow to return two data frames simultaneously.
  
  #' @return test/train data
  
  
    n_row = nrow(data)
    total_row = size * n_row
    train_sample <- 1: total_row
    if (train == TRUE) {
        return (data[train_sample, ])
    } else {
        return (data[-train_sample, ])
    }
}
```

Getting data
```{r}
data_train <- create_train_test(df, 0.8, train = TRUE)
data_test <- create_train_test(df, 0.8, train = FALSE)
dim(data_train)
```

Seeing propotion of data
```{r}
prop.table(table(data_train$income))
```
```{r}
prop.table(table(data_test$income))
```

fITTING
```{r}
fit <- rpart(income~., data = data_train, method = 'class')
rpart.plot(fit, extra = 106)
```

```{r}
predict_unseen <-predict(fit, data_test, type = 'class')
```

```{r}
table_mat <- table(data_test$income, predict_unseen)
table_mat
```

Evaluations
```{r}
accuracy_Test <- sum(diag(table_mat)) / sum(table_mat)
print(paste('Accuracy for test', accuracy_Test))
```
















